---
title: "Sentiment Analysis"
description: |
  This is the sentiment update I did in ESM 244 for a paper that I cite super often!
output: 
  distill::distill_article:
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
```

## Overview:
Here, I perform text wrangling and sentiment analysis with Oliver et al. (2019): Longer and more frequent marine heatwaves over the past century. This paper is meaningful to me because I study the impacts of marine heatwaves on fishery species in the Santa Barbara Channel. 

**Data citation:** Oliver, E. C. J., Donat, M. G., Burrows, M. T., Moore, P. J., Smale, D. A., Alexander, L. V., Benthuysen, J. A., Feng, M., Sen Gupta, A., Hobday, A. J., Holbrook, N. J., Perkins-Kirkpatrick, S. E., Scannell, H. A., Straub, S. C., & Wernberg, T. (2018). Longer and more frequent marine heatwaves over the past century. Nature Communications, 9(1), 1–12.


```{r, cache = TRUE}
### Get Oliver et al. 2019
oliver_text <- pdf_text(here::here('Oliver et al 2019.pdf'))
```


```{r, results='hide'}
### Get it into a data frame
oliver_lines <- data.frame(oliver_text) %>% 
  mutate(page = 1:n()) %>%
  mutate(text_full = str_split(oliver_text, pattern = '\\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 
```


```{r, results='hide'}
### Get some word counts by page
oliver_words <- oliver_lines %>% 
  unnest_tokens(word, text_full) %>% 
  select(-oliver_text)

oliver_wordcount <- oliver_words %>% 
  count(page, word)

# remove stop words
head(stop_words)
 
oliver_words_clean <- oliver_words %>% 
  anti_join(stop_words, by = 'word')

# recount
nonstop_counts <- oliver_words_clean %>% 
  count(page, word)

```


```{r, results='hide', fig.show='hide'}
### Find top 5 words on each page
top_5_words <- nonstop_counts %>% 
  group_by(page) %>% 
  arrange(-n) %>% 
  slice(1:5) %>%
  ungroup()
 
# graph it 
ggplot(data = top_5_words, aes(x = n, y = word)) +
  geom_col(fill = "blue") +
  facet_wrap(~page, scales = "free")

```

### Word Cloud 

```{r}
top100 <- nonstop_counts %>% 
  arrange(-n) %>% 
  slice(1:100)

cloud <- ggplot(data = top100, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "diamond") +
  scale_size_area(max_size = 6) +
  scale_color_gradientn(colors = c("darkgreen","blue","purple")) +
  theme_minimal()

cloud

```

**Figure 1:** The top 100 most frequently used words in Oliver et al. (2019).

### Sentiment analysis

```{r, results='hide'}
get_sentiments(lexicon = "afinn")
 
oliver_afinn <- oliver_words_clean %>% 
  inner_join(get_sentiments("afinn"), by = 'word')

# Get counts 
afinn_counts <- oliver_afinn %>% 
  count(page, value)
 
# Plot them: 
# ggplot(data = afinn_counts, aes(x = value, y = n)) +
#   geom_col() +
#   facet_wrap(~page)
 
# Find the mean afinn score by chapter: 
afinn_means <- oliver_afinn %>% 
  group_by(page) %>% 
  summarize(mean_afinn = mean(value))
 
ggplot(data = afinn_means, 
       aes(x = fct_rev(factor(page)),
           y = mean_afinn, fill = page)) +
  geom_col() +
  labs(x = "Mean afinn score", y = "Page number") +
  coord_flip() +
  theme(legend.position="none") +
  theme_minimal() +
  theme(legend.position = 'none')

```

**Figure 2:** Sentiment analysis with "afinn" by page number for Oliver et al. (2019)