---
title: "Climatic variables"
description: |
  This is an assignment I did in ESM 244 on various climatic variables.
output: 
  distill::distill_article:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
 
library(tidyverse)
library(here)
library(ggfortify) # For PCA biplot
library(patchwork)

```

### Introduction:
Here I explore miscellaneous environmental and climatic variables (at the country-level) and run principal component analysis (PCA) on the subset data. I select the variables elevation, mean annual rainfal, mean annual temperature range, mean annual temperature, mean wind, and mean cloudiness. The raw data were compiled and provided by @zander_venter on Kaggle; data is all acquired through Google Earth Engine (https://earthengine.google.com/). 

### Data wrangling

```{r}
# Read in the data & create a subset
env_vars_clean <- read_csv(here("world_env_vars.csv")) %>% 
  drop_na() %>%  # Dropped NA values
  select(Country, elevation, rain_mean_annual, temp_annual_range, temp_mean_annual, wind, cloudiness) # Selected 6 variables
```

### Principal Component Analysis 

```{r}
env_vars_pca <- env_vars_clean %>% 
  select(elevation:cloudiness) %>% 
  scale() %>%
  prcomp()

# see the loadings (weighing for each principal component)
env_vars_pca$rotation
```

```{r}
autoplot(env_vars_pca,
         data = env_vars_clean,
         loadings = TRUE, 
         loadings.label = TRUE,
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label.vjust = -0.5) +
  #scale_color_manual() +
  theme_minimal()

```

**Figure 1:** PCA plot of six environmental and climactic variables: elevation, mean annual rainfal, mean annual temperature range, mean annual temperature, mean wind, and mean cloudiness. 


### Summary 
- Mean annual rainfall and cloudiness appear to be positively correlated with each other.
- Elevation and mean annual temperature are negatively correlated. 
- PC1 explains 45.41% of the variance.
- PC2 explains 22.71% of the variance.

